---
title: "EDA"
author: "Margaret Reed"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r load-pkgs}
library(tidyverse)
library(lubridate)
library(prophet)
library(MLmetrics)
```


```{r load-data}
data <- read_csv("../raw_data/weekly_usage.csv")
```
## Exploratory data analysis

### data cleaning

To clean the data, all that was really required was creating a date object with the year and week variables.

```{r clean-data}
data <- data %>%
  mutate(
    date = lubridate::parse_date_time(paste(year, week, 1, sep="/"),'Y/W/w')
  ) %>%
  select(date, account, value, t)
```

### plotting the data

First I plotted the data as it was. 


```{r plot-data}
data %>%
   ggplot(aes(x = date, y = value)) +
   geom_point() +
  theme_minimal()
```


It looks pretty messy but there are clearly several different trends within the data.

### finding trends

To get a better look at these trends I decided to look at the 5 top consumers of electricity among the different accounts

```{r explore-data}
top_5_accounts <- data %>%
  group_by(account) %>%
  summarize(
    total_val = sum(value)
  ) %>%
  arrange(desc(total_val)) %>%
  slice(1:5) %>%
  pull(account)

data %>%
  filter(
    account %in% top_5_accounts
  ) %>%
   ggplot(aes(x = date, y = value, color = account)) +
   geom_line() +
  theme_minimal()
```


Clearly accounts MT_362, _196, _279, _370, and _208 are the top users and have fairly defined trends with seasonal components. 


First I will look at account MT_362

```{r modeling}
MT_362 <- data %>%
  filter(account == "MT_362") %>%
  select(date, value) %>%
  rename(ds = date, y = value)

n <- nrow(MT_362)
N_train <- round(0.7*n, 0)
N_validation <- round(0.2*n, 0)
N_test <- n - N_train - N_validation

train <- MT_362 %>%
  slice(1:N_train)

validation <- MT_362 %>%
  anti_join(train) %>%
  slice(1: N_validation)

test <- MT_362 %>%
  anti_join(train) %>%
  anti_join(validation)

 m <- prophet(rbind(train, validation))
```


```{r make-future}
future <- make_future_dataframe(m, periods = N_test, freq = "week")
```

```{r forecast}
forecast <- predict(m, future)
```

```{r}
MT_362 %>%
  ggplot(aes(x = ds, y = y)) +
  geom_line() +
  # geom_smooth(se = F) +
  theme_minimal()
```


```{r}
pred_test <- forecast %>%
  slice_tail(n=N_test) %>%
  pull(yhat)

true_test <- test %>%
  pull(y)
```


```{r}
full_join(MT_362, forecast) %>%
  ggplot() +
  geom_line(aes(x = ymd(ds), y = y), color = "black") +
  geom_line(aes(x = ymd(ds), y = yhat), color = "blue") +
  theme_minimal() +
  geom_line(data = MT_362%>%slice_head(n=N_train), aes(x = ymd(ds), y = y), color = "grey")
```
```{r}
s <- round(N_test/3, 0)

MAPE(pred_test, true_test)

MAPE(pred_test[1:s], true_test[1:s])
MAPE(pred_test[(s+1):(2*s)], true_test[(s+1):(2*s)])
MAPE(pred_test[(2*s+1):(3*s)], true_test[(2*s+1):(3*s)])
```

```{r}
tibble(
 first = map2_dbl(pred_test[1:s], true_test[1:s], MAPE),
 second = map2_dbl(pred_test[(s+1):(2*s)], true_test[(s+1):(2*s)], MAPE),
third = map2_dbl(pred_test[(2*s+1):(3*s)], true_test[(2*s+1):(3*s)], MAPE)
) %>%
  pivot_longer(
    cols = everything(),
    names_to = "segment",
    values_to = "error"
  ) %>%
  ggplot(aes(x = error, y = segment)) +
  geom_boxplot() +
  scale_x_continuous(limits = c(0, 1)) +
  labs(
    x = "Error (MAPE)",
    y = "Third of testing data",
    title = "Boxplots of errors"
  ) +
  theme_minimal()
```

